---
title: "Assignment 5: Data Visualization"
author: "Shivani Kuckreja"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

#install.packages("formatR")

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data
Visualization 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each 
instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a 
single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in 
Sakai. Add your last name into the file name 
(e.g., "Fay_A05_DataVisualization.Rmd") prior to submission.

The completed exercise is due on Monday, February 14 at 7:00 pm.

## Set up your session 
1. Set up your session. Verify your working directory and load the tidyverse 
and cowplot packages. Upload the NTL-LTER processed data files for nutrients 
and chemistry/physics for Peter and Paul Lakes (use the tidy 
[`NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv`] version) and 
the processed data file for the Niwot Ridge litter dataset (use the [`NEON_NIWO_Litter_mass_trap_Processed.csv`] version). 

2. Make sure R is reading dates as date format; if not change the format to date.

```{r}
#1
library(formatR)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE, echo=TRUE)

getwd()
library(tidyverse)
#install.packages("cowplot")
library(cowplot)


PeterPaul_chem_nutrients <- 
  read.csv(paste0("./Data/Processed/",
             "NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv"))

NW_litter <-
  read.csv("./Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv")

#2 
class(PeterPaul_chem_nutrients$sampledate) #originally a character
# Change date to date object
PeterPaul_chem_nutrients$sampledate <- ymd(PeterPaul_chem_nutrients$sampledate)

class(NW_litter$collectDate) #originally a character
# Change date to date object
NW_litter$collectDate <- ymd(NW_litter$collectDate)

```
## Define your theme
3. Build a theme and set it as your default theme.

```{r, fig.height = 3, fig.width = 4, warning = FALSE}
#3

mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "blue"), 
legend.position = "top") #alternative: legend.position + legend.justification

theme_set(mytheme)
```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best 
practices for data visualization. Ensure your theme, color palettes, axes, 
and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus (`tp_ug`) by phosphate (`po4`), with 
separate aesthetics for Peter and Paul lakes. Add a line of best fit and 
color it black. Adjust your axes to hide extreme values (hint: change the 
limits using `xlim()` and `ylim()`).

```{r}
#4
NTLLTER_Plot <- ggplot(PeterPaul.chem.nutrients, aes(x = tp_ug, y = po4, 
                                                      color=lakename)) +
  geom_point() +
  geom_smooth(method = lm, color = "grey") +
  ylab('Po4') +
  xlim(0, 200) +
  ylim(0, 20) 
print(NTLLTER_Plot)

```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, 
(b) TP, and (c) TN, with month as the x axis and lake as a color 
aesthetic. Then, create a cowplot that combines the three graphs. 
Make sure that only one legend is present and that graph axes are aligned.

```{r}
#5

NTLLTER_BoxPlot_Temp <- ggplot(PeterPaul.chem.nutrients, 
                               aes(x = month, y = temperature_C, 
                                   color=lakename)) +
  geom_boxplot() +
  scale_y_continuous(expand = c(0, 0)) +
  theme(legend.position = "right")
print(NTLLTER_BoxPlot_Temp)

NTLLTER_BoxPlot_TP <- ggplot(PeterPaul.chem.nutrients, aes(x = month, 
                                                          y = tp_ug, 
                                                          color=lakename)) +
  geom_boxplot() +
  scale_y_continuous(expand = c(0, 0)) +
  theme(legend.position = "right")
print(NTLLTER_BoxPlot_TP)


NTLLTER_BoxPlot_TN <- ggplot(PeterPaul.chem.nutrients, 
                             aes(x = month, y = tn_ug, color=lakename)) +
  geom_boxplot() +
  scale_y_continuous(expand = c(0, 0)) +
  theme(legend.position = "right")
print(NTLLTER_BoxPlot_TN)

#install.packages("cowplot") 
library(cowplot)
plot_grid(NTLLTER_BoxPlot_Temp, NTLLTER_BoxPlot_TP, 
          NTLLTER_BoxPlot_TN, nrow = 3, align = 'h', rel_heights = c(1.25, 1)) 

```

Question: What do you observe about the variables of interest over 
seasons and between lakes? 

> Answer: 

tp_ug: There are a lot of outliers in this dataset, especially 
compared to the Temperature plots. There is low variability in 
the data for Paul Lake.  

tn_ug: There are a lot of outliers in this dataset, and Peter Lake 
seems to have significantly more outliers (of higher value) than does Paul Lake. 

temperature: Temperature over seasons and between lakes appears 
to vary very little. 

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying 
only the "Needles" functional group. Plot the dry mass of needle litter 
by date and separate by NLCD class with a color aesthetic. (no need to 
adjust the name of each land use)

7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated 
into three facets rather than separated by color. 

```{r}
#6
NW_litter_filter <- NW_litter %>% 
  filter(functionalGroup == "Needles") 

ggplot(filter(NW_litter_filter)) +
  geom_point(aes(x = collectDate, y = dryMass,color=nlcdClass), size=1)


#7
NW_litter_filter <- NW_litter %>% 
  filter(functionalGroup == "Needles") 

ggplot(filter(NW_litter_filter)) +
  geom_point(aes(x = collectDate, y = dryMass), color="pink", size=1) +
facet_wrap('nlcdClass') 


```
Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: Plot 6 is more effective because it allows easier comparison across 
NLCD classes. Plot 7 only allows easier comparison by year within each NLCD class.  
