---
title: 'Assignment 6: GLMs (Linear Regressios, ANOVA, & t-tests)'
author: "Shivani Kuckreja"
output:
  pdf_document: default
  word_document: default
geometry: margin=2.54cm
editor_options:
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on 
generalized linear models. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each 
instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into 
a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox 
in Sakai. Add your last name into the file name (e.g., "Fay_A06_GLMs.Rmd") 
prior to submission.

The completed exercise is due on Monday, February 28 at 7:00 pm.

## Set up your session 
1. Set up your session. Check your working directory. Load the tidyverse, 
agricolae and other needed packages. Import the *raw* NTL-LTER raw data file 
for chemistry/physics (`NTL-LTER_Lake_ChemistryPhysics_Raw.csv`). Set date 
columns to date objects.

2. Build a ggplot theme and set it as your default theme.

```{r setup}
#1
setwd("/home/guest/R/Environmental_Data_Analytics_2022")
clelibrary(tidyverse)
#install.packages("agricolae")
library(agricolae)
ChemPhysRawData <- 
  read.csv("./Data/Raw/NTL-LTER_Lake_ChemistryPhysics_Raw.csv", 
           stringsAsFactors = TRUE)

class(ChemPhysRawData$sampledate) #originally a factor
# Change date to date object
ChemPhysRawData$sampledate <- mdy(ChemPhysRawData$sampledate)

#2
mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "blue"), legend.position = "top") 

theme_set(mytheme)

```

## Simple regression
Our first research question is: Does mean lake temperature recorded during July
change with depth across all lakes?

3. State the null and alternative hypotheses for this question:
> Answer:
H0: There is no relationship between mean lake temperature in July and depth
across lakes (the slope is equal to zero)
Ha: The slope is not equal to zero (For example: Mean lake temperature 
in July decreases as depth increases across lakes.)


4.  Wrangle your NTL-LTER dataset with a pipe function so that the records 
meet the following criteria: 
 * Only dates in July. 
 * Only the columns: `lakename`, `year4`, `daynum`, `depth`, `temperature_C`
 * Only complete cases (i.e., remove NAs)

5. Visualize the relationship among the two continuous variables with a 
scatter plot of temperature by depth. Add a smoothed line showing the 
linear model, and limit temperature values from 0 to 35 °C. Make this plot 
look pretty and easy to read.

```{r scatterplot}
#4

ChemPhysTempDepth <- ChemPhysRawData %>% 
mutate(Month = month(sampledate)) %>%
filter(Month %in% c("7")) %>%
select(lakename, year4, daynum, depth, temperature_C) %>%
filter(!is.na(lakename) & !is.na(year4) & !is.na(daynum) &
         !is.na(depth) & !is.na(temperature_C))

#5

TempbyDepth <- 
  ggplot(ChemPhysTempDepth, aes(x = depth, y = temperature_C)) +
  ylim(0, 35) +
  geom_point() +
  geom_smooth(method = lm, color = "grey") 
print(TempbyDepth) 

```


6. Interpret the figure. What does it suggest with regards to the response 
of temperature to depth? Do the distribution of points suggest about anything 
about the linearity of this trend?

> Answer: 
The figure suggests that, as depth increases, temperature decreases. It
is worth noting that there is increased variability in data points collected in 
shallower water vs. in deeper water, which undermines the linearity of 
the trend. 


7. Perform a linear regression to test the relationship and display the results

```{r linear.regression}
#7

Temp.Depth.Regression <- 
  lm(ChemPhysTempDepth$temperature_C ~ ChemPhysTempDepth$depth)
summary(Temp.Depth.Regression)

```


8. Interpret your model results in words. Include how much of the variability 
in temperature is explained by changes in depth, the degrees of freedom on 
which this finding is based, and the statistical significance of the result. 
Also mention how much temperature is predicted to change for every 1m 
change in depth. 

> Answer:
73.87 % of the variability in temperature is explained by changes in depth. 
This finding is based upon 9726 degrees of freedom. 
This result is statistically significant, as the p value is 2e-16 which is
less than 0.05. 
For every 1m change in depth, temperature is expected to change by about 2 
degrees Celcius. 

---

## Multiple regression
Let's tackle a similar question from a different approach. Here, we want 
to explore what might the best set of predictors for lake temperature 
in July across the monitoring period at the North Temperate Lakes LTER. 


9. Run an AIC to determine what set of explanatory variables 
(year4, daynum, depth) is best suited to predict temperature.

10. Run a multiple regression on the recommended set of variables. 

```{r temperature.model}
#9

TPAIC <- lm(data = ChemPhysTempDepth, temperature_C ~ year4 + daynum + depth)
#Choose a model by AIC in a Stepwise Algorithm
step(TPAIC)


#10
TPmodel <- lm(data = ChemPhysTempDepth, temperature_C ~ year4 + daynum + depth)
summary(TPmodel)

```

11. What is the final set of explanatory variables that the AIC method 
suggests we use to predict temperature in our multiple regression? 
How much of the observed variance does this model explain? Is this an 
improvement over the model using only depth as the explanatory variable?

> Answer: 
The final set of explanatory variables that the AIC method 
suggests we use to predict temperature in our multiple regression consists 
of year4, daynum, and depth. This model explains 74.12% of the observed 
variance, which is an improvement over the model using only depth as the 
explanatory variable, as that R-squared was only 73.87. 


---
## Analysis of Variance

12. Now we want to see whether the different lakes have, on average, different 
temperatures in the month of July. Run an ANOVA test to complete this analysis. 
(No need to test assumptions of normality or similar variances.) 
Create two sets of models: one expressed as an ANOVA models and another 
expressed as a linear model (as done in our lessons).

```{r anova.model}

#12

# Format ANOVA as aov
LakeTempJuly.Anova <- aov(data = ChemPhysTempDepth, temperature_C ~ lakename)
summary(LakeTempJuly.Anova)

# Format ANOVA as lm
LakeTempJuly.Anova2 <- lm(data = ChemPhysTempDepth, temperature_C ~ lakename)
summary(LakeTempJuly.Anova2)

# Checking model fit and assumptions
# ANOVA is robust against departures from normality.
plot(LakeTempJuly.Anova2)

```

13. Is there a significant difference in mean temperature among the lakes? 
Report your findings. 

> Answer: 
Yes, there is a significant difference in mean temperature among the lakes, as
the p-value is around 2e-16, which is less than 0.05. 


14. Create a graph that depicts temperature by depth, with a separate color 
for each lake. Add a geom_smooth (method = "lm", se = FALSE) for each lake. 
Make your points 50 % transparent. Adjust your y axis limits to go 
from 0 to 35 degrees. Clean up your graph to make it pretty. 

```{r scatterplot.2}
#14.
TempbyDepthByLake <- 
  ggplot(ChemPhysTempDepth, aes(x = depth, y = temperature_C, color=lakename)) +
  ylim(0, 35) +
  geom_point(alpha = 0.5) +
  geom_smooth (method = "lm", se = FALSE) 
print(TempbyDepthByLake)

```

15. Use the Tukey's HSD test to determine which lakes have different means.

```{r tukey.test}

#15
# Post-hoc test
# TukeyHSD() computes Tukey Honest Significant Differences
TukeyHSD(LakeTempJuly.Anova)

# Extract groupings for pairwise relationships
LakeTempJulyGroups <- HSD.test(LakeTempJuly.Anova, "lakename", group = TRUE)
LakeTempJulyGroups
```

16.From the findings above, which lakes have the same mean temperature, 
statistically speaking, as Peter Lake? Does any lake have a mean temperature 
that is statistically distinct from all the other lakes?

>Answer: 
From the findings above, the following lakes have the same mean temperature, 
statistically speaking, as Peter Lake: Central Long Lake, Crampton Lake, 
Ward Lake, Paul Lake,East Long Lake, Hummingbird Lake, Tuesday Lake,
and West Long Lake. No lake has a mean temperature that is statistically 
distinct from all the other lakes. 

17. If we were just looking at Peter Lake and Paul Lake. What's another test 
we might explore to see whether they have distinct mean temperatures? 

>Answer: 
Bartlett's test, which tests for variance between groups. 

