---
title: "Assignment 09: Data Scraping"
author: "Shivani Kuckreja"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

# Total points:

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on 
data scraping. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each 
instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a 
single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in 
Sakai. Add your last name into the file name (e.g., "Fay_09_Data_Scraping.Rmd") 
prior to submission.


## Set up 
1. Set up your session:

* Check your working directory
* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Set your ggplot theme

```{r, message = FALSE}
#1
getwd()
library(tidyverse)
library(lubridate)

#install.packages("rvest")
library(rvest)

# Set theme
mytheme <- theme_classic() +
  theme(axis.text = element_text(color = "blue"), 
        legend.position = "top")
theme_set(mytheme)

```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2019 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Change the date from 2021 to 2020 in the upper right corner.
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: 
 <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2020>

Indicate this website as the as the URL to be scraped. 
(In other words, read the contents into an
`rvest` webpage object.)

```{r set.the.scraping.website}
#2
#Fetch the web resources from the URL
webpage <- 
read_html('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2020')
webpage

```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PWSID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Maximum Daily Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to four 
separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third 
"Municipality", and the last should be a vector of 12 numeric values, with 
the first value being 36.0100.

```{r scrape.the.data}
#3
water.system.name <- webpage %>% 
  html_nodes("div+ table tr:nth-child(1) td:nth-child(2)") %>% html_text()
pwsid <- webpage %>% 
  html_nodes("td tr:nth-child(1) td:nth-child(5)") %>% html_text()
ownership <- webpage %>% 
  html_nodes("div+ table tr:nth-child(2) td:nth-child(4)") %>% html_text()
max.withdrawals.mgd <- webpage %>% html_nodes("th~ td+ td") %>% html_text()

```


4. Convert your scraped data into a dataframe. This dataframe should have 
a column for each of the 4 variables scraped and a row for the month 
corresponding to the withdrawal data. Also add a Date column that 
includes your month and year in data format. (Feel free to add a Year 
column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data 
in order. You can overcome this by creating a month column in the same order 
the data are scraped: Jan, May, Sept, Feb, etc...

5. Plot the max daily withdrawals across the months for 2020

```{r create.a.dataframe.from.scraped.data}
#4
Month <- webpage %>% 
  html_nodes(".fancy-table:nth-child(31) tr+ tr th") %>% html_text()

Year <- 2020

the_df <- data.frame(
"Water System Name" = water.system.name,
"PWSID" = pwsid,
"Ownership" = ownership,
"Maximum Daily Use (MGD)" = max.withdrawals.mgd, 
"Month" = Month, 
"Year" = Year,
"Date" = my(paste(Month,"-",Year)
))


#5
#Plot 
ggplot(the_df, aes(x=Date,y=max.withdrawals.mgd)) + 
  geom_point() + 
  labs(title = paste("Max Daily Withdrawals Across the Months"),
       subtitle = 2020,
      # scale_x_date(date_breaks="Month"), # code to show all month labels 
      #on x-axis not working
       y="Max Withdrawal (mgd)",
       x="Date")

# geom_line not showing up which is why I am using geom_point

```

6. Note that the PWSID and the year appear in the web address for 
the page we scraped. Construct a function using your code above that 
can scrape data for any PWSID and year for which the NC DEQ has data. 
**Be sure to modify the code to reflect the year and site scraped**.

```{r construct.a.scraping.function}
#6. 
scrape.it <- function(Year, PWSID){
  the_url <- 
    read_html(paste0('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=',
    PWSID, '&year=',Year))
    print(the_url)


  water.system.name_revised <- the_url %>% 
  html_nodes("div+ table tr:nth-child(1) td:nth-child(2)") %>% html_text()
  pwsid_revised <- the_url %>% 
  html_nodes("td tr:nth-child(1) td:nth-child(5)") %>% html_text()
  ownership_revised <- the_url %>% 
  html_nodes("div+ table tr:nth-child(2) td:nth-child(4)") %>% html_text()
  max.withdrawals.mgd_revised <- the_url %>% html_nodes("th~ td+ td") %>% html_text()

  
max.withdrawls.year.new <- Year
  
dataframe <- data.frame("WaterSystemName" = as.character(water.system.name_revised), 
"PWSID" =  as.character(pwsid_revised),                 
"Ownership" =  as.character(ownership_revised),
"MaxWithdrawals" =  as.character(max.withdrawals.mgd_revised),
"Month" =  as.character(Month), 
"Years" = max.withdrawls.year.new)

dataframe <- dataframe %>% 
  mutate("Date" = my(paste(Month, "-", Year)))

return(dataframe)
}
```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2015

```{r fetch.and.plot.Durham.2015.data}
#7
#Similar to 5 except use function

withdrawals.2015 <- scrape.it (2015, '03-32-010')

ggplot(withdrawals.2015, aes(x=Date,y=MaxWithdrawals)) + 
  geom_point() + 
  labs(title = paste("Durham: Max Daily Withdrawals Across the Months"),
       subtitle = 2015,
      # scale_x_date(date_breaks="Month"), # code to show all month labels
      #on x-axis not working
       y="Max Withdrawal (mgd)",
       x="Date")

# geom_line not showing up which is why I am using geom_point
```

8. Use the function above to extract data for Asheville (PWSID = 01-11-010) 
in 2015. Combine this data with the Durham data collected above and create a 
plot that compares the Asheville to Durham's water withdrawals.

```{r fetch.and.plot.Durham.2015.data}
#8
Asheville.2015 <- scrape.it (2015, '01-11-010')

ggplot(Asheville.2015, aes(x=Date,y=MaxWithdrawals)) + 
  geom_point() + 
  labs(title = paste("Asheville: Max Daily Withdrawals Across the Months"),
       subtitle = 2015,
       y="Max Withdrawal (mgd)",
       x="Date")

# geom_line not showing up which is why I am using geom_point

#COMBINE
```


9. Use the code & function you created above to plot Asheville's max daily 
withdrawal by months for the years 2010 thru 2020.Add a smoothed line to 
the plot.

```{r}
#9
Year = seq(2010, 2020)
pwsid = '01-11-010'

Ashville2010to2020 <- Year %>%
  map(scrape.it)

```

>Question: Just by looking at the plot (i.e. not running statistics), 
does Asheville have a trend in water usage over time?